---
title: "Random Forest"
author: "Lisa Leung"
date: '2019-01-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load all libraries

```{r}
library(tidyverse)
library(caret)
library(randomForest)
library(quantregForest)
```
## Purpose

We will conduct random forest to predict the number of purchase made

```{r}
# Loading data
dat_User <- read_csv("BlackFriday-User.csv")
dat_User <- dat_User[,!colnames(dat_User) %in% c("X1", "User_ID")]
dat_User$Occupation <- as.factor(dat_User$Occupation) #converting to a factor


# Train and Test data
set.seed(10)
num <- round(nrow(dat_User)/2)
vec_Train <- sample(1:nrow(dat_User),size = num)

dat_Train <- dat_User[vec_Train,]
dat_Test <- dat_User[-vec_Train,]
```

## Multiple Regression

Before starting random forest, I want to use a simple multiple regression to understand simple correlations among the data set

```{r}
lm_multiple <- lm(sum_Purchase~Gender+Age+Occupation+City_Category+Stay_In_Current_City_Years + Marital_Status, data=dat_Train)
summary(lm_multiple)
```
According to a summary of the multiple linear regression, we see that gender, cities, and occupations play a huge role in total purchases made. I'd like to also see all three cities and check if there are any other differences among the cities

### Checking out regressions separated by cities
```{r}
dat_A <- dat_Train[dat_Train$City_Category=="A",]
lm_multiple_A <- lm(sum_Purchase~Gender+Age+Occupation+Stay_In_Current_City_Years + Marital_Status, data=dat_A)
summary(lm_multiple_A)

dat_B <- dat_Train[dat_Train$City_Category=="B",]
lm_multiple_B <- lm(sum_Purchase~Gender+Age+Occupation+Stay_In_Current_City_Years + Marital_Status, data=dat_B)
summary(lm_multiple_B)

dat_C <- dat_Train[dat_Train$City_Category=="C",]
lm_multiple_C <- lm(sum_Purchase~Gender+Age+Occupation+Stay_In_Current_City_Years + Marital_Status, data=dat_C)
summary(lm_multiple_C)
```

## Let's start the random forest!

```{r}
seed <- 10
set.seed(seed)

# Setting parameters on mtry tuning
control <- trainControl(method="repeatedcv", number=10, repeats=3) # using 10-fold, repeating 3 times from caret
metric <- "RMSE" #A string that specifies what summary metric will be used to select the optimal model. By default, possible values are "RMSE" and "Rsquared" for regression and "Accuracy" and "Kappa" for classification. 
mtry <- sqrt(ncol(dat_Train))
tunegrid <- expand.grid(.mtry=mtry) # Looking at the randomly selected predictors

# Running the random forest
# Note that I am doing a regression random forest so I'l be using quantile random forest
rf_default <- train(sum_Purchase~Gender+Age+Occupation+City_Category+Stay_In_Current_City_Years + Marital_Status, data=dat_Train, method = 'qrf', metric=metric, tuneGrid=tunegrid, trControl=control) #using quantile random forest to train. Tuning parameters uses mytry (randomly selected predictors)
print(rf_default)
```

Rsquared 

